Visualizing Functional Data

### 2017-02-08 sections/section0d.tex
more fleshing of section 0d bi/multi

The bar graph is arguable the simplest way to visualize data that is all of the
same type drawn from a sample at random. In this visualization technique, the
heights of the bar are representative of the values, the x values are each
invidual observation and the positioning of the x values is somewhat
arbitrary. \cite{TufteVisualDIsplay, Friendly, Playfair}. Since bar
graphs become somewhat useless as the number of samples increase, it is common
to group (bin) the data and then count the number of elements in each
been. Both this binning and counting and the subsequent bar chart of the
frequency are referred to as histograms \cite{ioannidis2003}. While histograms
give an approximation of the distribution of the data, sometimes the bin widths
can be too wide and thereby fail to catch naunces in the data such as
bimodality. A way to counteract this is to build histograms with very thin
buckets, which is one of the less sophisticated methods of doing kernal density
estimation. Other methods of approximating the distribution include nearest
neighbor estimation or the  moving windowed average density trace method \cite{chambers1983}.

@@

\subsubsection{Dependent variables}
%%scatter segue into bagplot and HDR?
Arguably the simplest and most well known method of visualizing two
variables against each other is the scatterplot\cite{friendly2005}. In
a scatter plot, one variable is mapped to the x coordinate and a
second variable is mapped to the y coordinate,

@@

\subsection{Multivariate Observations}

%%bubble plot? 
%% large scale box: letter, spark

%% radial, parallel coordinates, factor plotl
%%high d-PCA, ICA, markov

>>>>

The bar graph is arguably the simplest way to visualize data that is all of the
same type drawn from a sample at random. In this visualization technique, the
heights of the bar are representative of the values, the x values are each
invidual observation and the positioning of the x values is somewhat
arbitrary. \cite{TufteVisualDIsplay, Friendly, Playfair}. Since bar
graphs become somewhat useless as the number of samples increase, it is common
to group (bin) the data and then count the number of elements in each
been. Both this binning and counting and the subsequent bar chart of the
frequency are referred to as histograms \cite{ioannidis2003}. While histograms
give an approximation of the distribution of the data, sometimes the bin widths
can be too wide and thereby fail to catch naunces in the data such as
bimodality. A way to counteract this is to build histograms with very thin
buckets, which is one of the less sophisticated methods of doing kernal density
estimation. Other methods of approximating the distribution include nearest
neighbor estimation or the  moving windowed average density trace method \cite{chambers1983}.

@@

\subsubsection{Dependent variables}
%%scatter segue into bagplot and HDR?
Arguably the simplest and most well known method of visualizing two
variables against each other is the scatterplot\cite{friendly2005}. In
a scatter plot, one variable is mapped to the x coordinate and a
second variable is mapped to the y coordinate, and the distribution of the
points in the x/y plane indicates whether the x and y variables are related to
each other. While this plot gives a sense of direction, it can fail for larger
datesets (% insert citations for scatter for larger datasets) and doesn't
capture the uncertainity of the relationship between the x and y variables. To
display the distributional variance, the bagplot and bivariate HDR plot were
created. These plots work in X/Y/Z - bag/box papers tomorrow morning

@@

\subsection{Multivariate Observations}
If the variables are independent, then small multiples (insert citation) or
visualizations for a smaller number of variables can be used to visualize
them. But this technique is limited to a somewhat small number of variables
(maybe 20 (find paper)) before the number of multiples is overwhelming and
impedes decision making. Instead, there are some techniques to look at a large
set of independent variables at once, such as ....

%% large scale box: letter, spark

%%bubble plot
%% radial, parallel coordinates, factor plotl
%%high d-PCA, ICA, markov
